A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-05-07 FMCAD2021/paper.tex
abstract and references

\title{A Decision Procedure for String Constraints with {\parseInt} Function and Flat Regular Constraints}
\author{ }
\institute{ } 
\begin{document}
\maketitle{}

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion function, which appears in almost all scripting languages.
    However, their approach supports only a specific flattening pattern and leaves the support of the general class as an open problem.
    In this paper, we bridge the gap and propose a complete flattening approach for the string-integer conversion, based upon a decision procedure for solving the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions).
    Besides the theoretical results, we also develop various optimizations and implement the decision procedure.
    Moreover, we evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as generated randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer constraints as well as linear-exponential constraints, in both precision and efficiency.
\keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}

@@

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov}. 
Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
by \cite{semenov}.
Particularly, in \cite{Francoise} Francoise gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Francoise's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

In the next section, within Francoise's framework, we give a revised algorithm
 to the decision problem of {\paexp} with 
both subtle improvements and crucial corrections, that 
provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
Theorem~\ref{thm:string-parInt}.

\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
Based on Francoise's work \cite{Francoise}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
that is, $\exists x.\theta(x,\bar{y})$. 
For eliminating quantifiers in arbitrary formula, 
we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.

>>>>

\title{A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints}
\author{ }
\institute{ } 
\begin{document}
\maketitle{}

\begin{abstract}
    String constraint solving is the core of various testing and verification approaches for scripting languages. 
    Among algorithms for solving string constraints, flattening is a well-known approach that is particularly useful in handling satisfiable instances.
    In the PLDI 2020 paper of Abdulla et al., the authors extended the flattening approach to support the string-integer conversion, which is an important function appearing in almost all scripting languages.
    However, their approach supports only a special flattening pattern and leaves the support of the general flat regular constraints as an open problem.
    In this paper, we fill the gap and propose a complete flattening approach for the string-integer conversion. The approach is built upon a decision procedure for the linear-exponential arithmetic constraints (namely, the extension of Presburger arithmetic with exponential functions).
    Besides the theoretical results, we also develop various optimizations and provide an implementation.
    We evaluate the performance of the implementation on the benchmarks that are generated from the string hash functions as well as randomly.
    The experimental results show that our implementation outperforms the state-of-the-art solvers for string-integer conversion constraints as well as linear-exponential constraints, in both precision and efficiency.
\keywords{ String Constraints \and Decidability \and Exponential Functions \and Quantifier Elimination.}

@@

Semenov proved that for any function $f$ that is compatible with addition, 
theory $(\mathbb{N},+,f)$ admits QE  and thus the theory is complete and decidable \cite{Semenov84}. 
Exponential functions are compatible with addition functions, and therefore {\paexp} is decidable 
by \cite{Semenov84}.
Particularly, in \cite{Point86} Point gave a detailed QE algorithm for $(\mathbb{N},+,2^x)$, where $f(x)=2^x$.
Unfortunately, Point's algorithm is flawed, one problem 
 lies in the discussion to eliminate exponential terms. 
% making the algorithm fail to handle "$a_0<0$" case.
For example, after  eliminating variable $x$ in formula $\exists x. y\le 2^x \wedge x\le 10$,
the expected result should be $y\le 2^{10} =1024$,
but Francoise's algorithm will return a formula equivalent to $y \le 2^6=64$.
% 这里好像还有另一个与Francoise一起的作者

In the next section, within Point's framework, we give a revised algorithm
 to the decision problem of {\paexp} with 
both subtle improvements and crucial corrections, that 
provides a crucial step to the decision procedure of string constraints with {\parseInt} function according to 
Theorem~\ref{thm:string-parInt}.

\section{Quantifier Elimination Algorithm for $(\mathbb{N},+,2^x)$}
Based on Point's work \cite{Point86}, we present a  revised QE algorithm for the formula of the form $Qx.\theta(x,\bar{y})$, 
where $Q$ is a quantifier and $\theta(x,\bar{y})$ is a quantifier-free formula. 
Since $\forall x. F = \neg \exists x. \neg F$, we further assume the quantifier $Q$ to be the existential quantifier, 
that is, $\exists x.\theta(x,\bar{y})$. 
For eliminating quantifiers in arbitrary formula, 
we apply the algorithm to the innermost quantified formula and repeat this procedure until all quantifiers are eliminated.
